---
title: "Geog533 Lab 8"
author: "Your name"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
n <- 24
residual <- 1.7
total <- 2.3
regression <- total - residual

df.regression <- 1
df.residual <- n-2
df.total <- n-1

mean.regression <- regression/df.regression
mean.residual <- 1.7 / df.residual
mean.total <- df.total/2.3
f1 <- mean.regression/mean.residual
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|   `r regression` |  `r df.regression`   | `r mean.regression`|        `r f1`    |
| Residual |       1.7       |  `r df.residual`     | `r mean.residual` |                |
| Total    |       2.3       |  `r df.total`        | `r mean.total`  |                  |



### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
a <- 0.46
b <- 0.19
x <- 50000/1000
p <- a + b*x
round(p,digits = 1)
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

when one unit increases in annual income, the weekly shopping trip frequency would accerate 0.19

### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
f.critical <- qf(0.95, df1 = 1, df2 = n-2)
if(f < f.critical){
  print("we accept the null hypothdsis: the regression coefficient is different from zero")
}else {
  print("we reject the null hypothesis: the regression coefficient is not different from zero")
}
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
R2 <- regression/total
R <- sqrt(R2)
R
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
snowfall <- c(36, 78, 11,45)
elevation <- c(400,800,200,675)
snowfall.m <- mean(snowfall)
elevation.m <- mean(elevation)
m <- lm(snowfall ~ elevation)

Stop <- sum((snowfall-snowfall.m)*(elevation-elevation.m)) 
Strail <- sum((elevation-elevation.m)^2 )
slope <- Stop/Strail
intercept <- snowfall.m-slope*elevation.m
slope
intercept
```

```{r}
summary(m)
```

```{r}
print("the intercept and the slope are not significant")
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
n <- 4
m <- lm(snowfall ~ elevation)
result <- summary(m)
top <- sum((result$residuals - mean(result$residuals))^2)
bot <- n-2
sde <- sqrt(top/bot)
sde
```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
A <- anova(m)
f <- A$`F value`[1]
f.critical <- qf(0.95, df1 = 1, df2 = n-2)
if (f > f.critical){
  print("we reject the hypothesis that the independent variables is equal to zero")
}else{
  print("we accept the hypothesis that the independent variables is equal to zero")
}
```


### Question 2(d)
Find the value of $r^2$.
```{r}
R2 <- result$r.squared
R2
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
library(knitr)
snowfall <- c(36, 78, 11,45)
predict <- m$fitted.values
residual <- m$residuals
residual
df1 <- data.frame(snowfall, predict, residual)
kable(df1)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
result <- anova(m)
result$Df[[1]]
result$Df[[2]]
result$`Sum Sq`[[1]]
result$`Sum Sq`[[2]]
total <- result$`Sum Sq`[[1]]+result$`Sum Sq`[[2]]
result$`Mean Sq`[[1]]
result$`Mean Sq`[[2]]
total.m <- total/2
result$`F value`[[1]]
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|2047.77 |1 | 2047.77 |16.17339  |
| Residual |253.2274   |2 |  126.6137 |                 |
| Total    |2301        |        3         |              |                 |
### Question 2(g)
Graph the data and the regression line.
```{r}
plot(snowfall ~ elevation)
abline(m, col = "blue", lwd = 2)
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df3 <- read.csv(file = url)
y1 <- df3$SalePrice
x1 <- df3$LotSize
m1 <- lm(y1 ~ x1)
m1
plot(y1 ~x1)
abline(m1, col = "red", lwd = 3)
print("y = 92984.25 + 6.443x")
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
ur2 = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df4 <- read.csv(file = ur2)
y2 <- df4$price
x2 <- df4$bedrooms
m2 <- lm(y2 ~ x2)
m2
plot(y2 ~x2)
abline(m2, col = "purple", lwd = 3)
print("y = 11254 + 11893x")
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
ur2 = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df4 <- read.csv(file = ur2)
y3 <- df4$price
x3 <- df4$bathrooms
m <- lm(y3 ~ x3)
m
plot(y3 ~x3)
abline(m, col = "pink", lwd = 3)
print("y = 32729 + 7113x")
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
library(leaflet)
m <- leaflet(width = "100%") %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=106.7828953, lat=26.622512, popup="The birthplace of Xiran He")
m  # Print the map
```

