---
title: "Geog533 Lab 10"
author: "Your  name"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
library(UScensus2010)
library(UScensus2010county)
library(UScensus2010tract)
data("new_york.county10")
shp <- new_york.county10
plot(shp)

```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
library(tmap)
qtm(shp)
qtm(shp,fill = "NAME10")
```


### Question 1(c)	
How many counties in New York State?
```{r}
library(UScensus2010county)
data("new_york.county10")
shp <- new_york.county10
df <- shp@data
number <- nrow(df)
number
print("there are 62 counties in New York State")
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
df$county[df$NAME10 == "Broome"]
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
pop <- df$P0010001
sm <- summary(pop)
total <- sum(pop)
sk <- skewness(pop)
min <- sm[[1]]
max <- sm[[6]]
mean <- sm[[4]]
median <- sm[[3]]
data.frame(total,min,max,mean,median,sk)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop)
boxplot(pop)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}
library(UScensus2010)
library(UScensus2010county)
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
data("new_york.tract10")
shp.t <- new_york.tract10
plot(shp.t)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df.t <- shp.t@data
pop.t <- df.t$P0010001
total.t <- sum(pop.t)
total.t
print("the total population is 19378102 ")
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
broome <- df.t$county == "007"
plot(shp.t[broome,])
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
Broome <- df.t[df.t$county == "007",]
Bpop <- Broome$P0050001
total2 <- sum(Bpop)
total2
print("the total population of Broome country is 200600")
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(Bpop)
boxplot(Bpop)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
broome5 <- Broome[,1:5]
broome5$ratio <- Broome$P0010001/total2
shp.B <- shp.t[broome,]
shp.B@data <-broome5 
df <- shp.B@data
writeOGR(shp.B,dsn = "output",layer = "shp.B",driver="ESRI Shapefile",overwrite_layer = TRUE)
shp <- readOGR(dsn = "output", layer = "shp.B")
plot(shp)
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r}
if(!require(ncdf4)) {install.packages("ncdf4")}
library(ncdf4)
library(raster)

```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
setwd("NDVI")
ndvi.nc <- brick("NDVI.nc")
ndvi.nc
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi.nc)
ncol(ndvi.nc)
ncell(ndvi.nc)
extent(ndvi.nc)
bbox(ndvi.nc)
res(ndvi.nc)
projection(ndvi.nc)
nbands(ndvi.nc)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
ndvimean <- mean(ndvi.nc) 
plot(ndvimean)
ndvimax <- max(ndvi.nc)
plot(ndvimax)
writeRaster(ndvimax,filename = "ndvimax")
writeRaster(ndvimean,filename = "ndvimean")
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
ndvi2001 <- ndvi.nc[[13:24]]
plot(ndvi2001)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r, warning=FALSE}
ndvi2001 <- ndvi.nc[[13:24]]
hist(ndvi2001)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
plot(ndvi.nc,7)
#values <- click(ndvi.nc, n=1, xy=TRUE) 
#values <- click(ndvi.nc, n=1, xy=FALSE)
values <- ndvi.nc[50,59]
plot(as.vector(values), type="b")
```

