---
title: "Sentiment Analysis"
author: "He Xiran"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Creating World Cloud
In this part I want to create a world cloud about Trump. According to this way I what words people will use to evaluate Trump.

### Authentication with OAuth
```{r, message=FALSE, warning=FALSE}
library(twitteR)
library(ROAuth)

# you must get the following information from the Twitter App you just created
my.consumer.key = "jGMlhMS8kBFdXd8eYZElRbzJH"
my.consumer.secret = "HMTxTdQ1IruiQT4PkeK0upLxKpsBBcql2gDIkUUOXfoIG0KYFl"
my.access.token = "917361228456095745-uJSuEQNSwGhrYhhycgX9i3UHRo73DnH"
my.access.token.secret = "RQFqWp5xI0gv1uzi41MfH5F2UWjLw3hhdVljrQOO4hxmP"

my_oauth <- setup_twitter_oauth(consumer_key = my.consumer.key, consumer_secret = my.consumer.secret, access_token = my.access.token, access_secret = my.access.token.secret)
```

### Extract Tweets	

```{r}
tweets <- searchTwitter("#Trump", n=1000, lang="en")
tweets.text <- sapply(tweets, function(x) x$getText())
```


### Clean Up Text	and remove stop words	

```{r, message=FALSE, warning=FALSE}
# Replace blank space (“rt”)
tweets.text <- gsub("rt", "", tweets.text)

# Replace @UserName
tweets.text <- gsub("@\\w+", "", tweets.text)

# Remove punctuation
tweets.text <- gsub("[[:punct:]]", "", tweets.text)

# Remove links
tweets.text <- gsub("http\\w+", "", tweets.text)

# Remove tabs
tweets.text <- gsub("[ |\t]{2,}", "", tweets.text)

# Remove blank spaces at the beginning
tweets.text <- gsub("^ ", "", tweets.text)

# Remove blank spaces at the end
tweets.text <- gsub(" $", "", tweets.text)
 
# #convert all text to lower case
tweets.text <- tolower(tweets.text)

if(!require(tm)) {install.packages("tm")}
library(tm)

#create corpus
tweets.text.corpus <- Corpus(VectorSource(tweets.text))

#clean up by removing stop words
tweets.text.corpus <- tm_map(tweets.text.corpus, function(x) removeWords(x,stopwords()))
```

### Generate Word Cloud	

```{r, message=FALSE, warning=FALSE}
if(!require(wordcloud)) {install.packages("wordcloud")}
library(wordcloud)

#generate wordcloud
wordcloud(tweets.text.corpus,min.freq = 2, scale=c(7,0.5),colors=brewer.pal(8, "Dark2"),  random.color= TRUE, random.order = FALSE, max.words = 150)
```

## Part 2. Sentiment Analysis

### Read Positive and Negative Words
 
```{r, message=FALSE, warning=FALSE}
setwd("C:/Users/Lei/Documents/First year of graduate/Statistics for Geography/Final project/Final project")
      neg = scan("negative-words.txt", what="character", comment.char=";")
      pos = scan("positive-words.txt", what="character", comment.char=";")
```

### Function for Scoring Tweets
Create a histogram and a boxplot of the population.
```{r}
score.sentiment = function(tweets, pos.words, neg.words)

{
scores = laply(tweets, function(tweet, pos.words, neg.words) {

tweet = gsub('https://','',tweet) # removes https://
tweet = gsub('http://','',tweet) # removes http://
tweet=gsub('[^[:graph:]]', ' ',tweet) ## removes graphic characters  #like emoticons 
tweet = gsub('[[:punct:]]', '', tweet) # removes punctuation 
tweet = gsub('[[:cntrl:]]', '', tweet) # removes control characters
tweet = gsub('\\d+', '', tweet) # removes numbers
tweet=str_replace_all(tweet,"[^[:graph:]]", " ") 
tweet = tolower(tweet) # makes all letters lowercase

word.list = str_split(tweet, '\\s+') # splits the tweets by word in a list
words = unlist(word.list) # turns the list into vector
pos.matches = match(words, pos.words) ## returns matching values for words from list 
neg.matches = match(words, neg.words)
pos.matches = !is.na(pos.matches) ## converts matching values to true of false
neg.matches = !is.na(neg.matches)
 
score = sum(pos.matches) - sum(neg.matches) # true and false are treated as 1 and 0 so they can be added
 
return(score)
 
}, pos.words, neg.words )
 
scores.df = data.frame(score=scores, text=tweets)
 
return(scores.df)
 
}
```


### Extract Tweets
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r, message=FALSE, warning=FALSE}
library(stringr)
library(plyr)
tweets = searchTwitter('Trump',n=500)
Tweets.text = laply(tweets,function(t)t$getText()) # gets text from Tweets
analysis = score.sentiment(Tweets.text, pos, neg) # calls sentiment function

```

### Calculate the proportion of netizen who have a positive attitude towards Trump	
Plot a map of New York census tracts using the plot function.
```{r, message=FALSE, warning=TRUE}
Tweets <- list()
Tweets.text <- list()
analysis <- list()
sup_ratio <- list()
weight_ratio <- list()

for(i in 1:15){
        Tweets[[i]] <- searchTwitter('Trump',n = 500,locale = 11/i/17)
        Tweets.text[[i]] = laply(Tweets[[i]],function(t)t$getText()) # gets text from Tweets
        analysis[[i]] = score.sentiment(Tweets.text[[i]], pos, neg) # calls sentiment function
        k <- 0
        cal <- analysis[[i]]$score
        for(j in 1:500){
          if(cal[j]>=1){
            k <- k+1
          } 
        }
        sup_ratio[i] <- k/500
}
ratio <- as.numeric(sup_ratio)
hist(ratio)
```

### Calculate the weight ratio of netizens' attitude

```{r, message=FALSE, warning=FALSE}
for(i in 16:30){
  Tweets[[i]] <- searchTwitter('Trump',n = 500,locale = 11/i/17)
  Tweets.text[[i]] = laply(Tweets[[i]],function(t)t$getText()) # gets text from Tweets
  analysis[[i]] = score.sentiment(Tweets.text[[i]], pos, neg) # calls sentiment function
  k <- 0
  cal <- analysis[[i]]$score
  for(j in 1:500){
    if(cal[j]>=1){
      k <- k+1
    } 
  }
  sup_ratio[i] <- k/500
}
for(i in 1:30){
  k <- 0
  g <- 0
  cal <- analysis[[i]]$score
  for(j in 1:500){
    if(cal[j]>=1){
      k <- k+cal[j]
    }
    if(cal[j]<0){
      g <- g+abs(cal[j])
    }
  }
  weight_ratio[i] <- k/(k+g)
}
weight_ratio <- as.numeric(weight_ratio)
hist(weight_ratio)
```

## Analysis correction
### The relationship between positive attitude and the approval rating of Trump

```{r}
setwd("C:/Users/Lei/Documents/First year of graduate/Statistics for Geography/Final project/Final project")
      neg1 = scan("R.txt", what="character", comment.char=";")
approve <- as.numeric(neg1)
ratio <- as.numeric(sup_ratio)


x <- approve
y <- ratio
plot(y ~ x)
m <- lm(y~x)
summary(m)
abline(m,col="red",lwd=2)
segments(x,fitted(m),x,y,col="blue",lty = "dashed")
```

```{r}
print("the relationship between approval rating of Trump and attitude of netizen is:y*attitude = 1.09922*approval rating + 0.13515")
```

### ### The relationship between weight positive attitude and the approval rating of Trump

```{r}
Wratio <- as.numeric(weight_ratio)

weight_x <- approve
weight_y <- Wratio
plot(weight_y ~ weight_x)
weight_m <- lm(weight_y~weight_x)
summary(weight_m)
abline(weight_m,col="blue",lwd=2)
segments(weight_x,fitted(weight_m),weight_x,weight_y,col="blue",lty = "dashed")
```

```{r}
print("the relationship between approval rating of Trump and weight positive attitude is:y*weight_attitude = 0.8124*approval rating + 0.46417")
```

