---
title: "Geog533 Lab 9"
author: "Your  name"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 9 (More on Regression) of the textbook pages 286-287. 


## Question 1

This is Exercise 7 in Chapter 9 of the Textbook [R].

The following results were obtained from a regression of $n = 14$ housing prices (in dollars) on median family income, size of house, and size of lot:

```{r, echo=FALSE}
library(knitr)
col1 <- c("Regression SS:","Residual SS:","Total SS:")
col2 <- c("4234","3487","")
col3 <- c("3","","")
col4 <- c("","","")
col5 <- c("","","")
df <- cbind(col1,col2,col3,col4,col5)
colnames(df) <- c("","Sum of squares","df","Mean square","F")
kable(df)
```

```{r, echo=FALSE}
library(knitr)
col1 <- c("Median family income","Size of house (sq.ft)","Size of lot (sq.ft)","Constant")
col2 <- c("1.57","23.4","-9.5","40,000")
col3 <- c("0.34","11.2","7.1","1000")
col4 <- c("1.3","2.9","11.3","")
df <- cbind(col1,col2,col3,col4)
colnames(df) <- c("","Coefficient (b)","Standard error (sb)","VIF")
kable(df)
```


### Question 1(a)
Fill in the blanks.

```{r}
library(knitr)
col1 <- c("Regression SS:","Residual SS:","Total SS:")
col2 <- c("4234","3487","7721")
col3 <- c("3","10","13")
col4 <- c("1411.333","348.7","593.923")
col5 <- c("4.0474","","")
df <- cbind(col1,col2,col3,col4,col5)
colnames(df) <- c("","Sum of squares"," df ","Mean square "," F")
kable(df)
```


### Question 1(b)
What is the value of $r^2$? 

```{r}
r2 <- 4234/7721
r2
```


### Question 1(c)
What is the standard error of the estimate? 

```{r}
se <- sqrt(348.7)
se
```


### Question 1(d)
Test the null hypothesis that $R^2 = 0$ by comparing the $F-statistic$ from the table with its critical value. 
```{r}
F <- 4.0474
f <- qf(0.95,3,10)
f
if(F<f){
  print("we accept null hypothesis that r2 = 0")
}else{
  print("we reject null hypothesis that r2 = 0")
}

```


### Question 1(e)
Are the coefficients in the direction you would hypothesize? If not, which coefficients are opposite in sign from what you would expect? 

```{r}
print("the size of lot is not in the direction I would hypothesize")
```


### Question 1(f)
Find the $t-statistics$ associated with each coefficient, and test the null hypotheses that the coefficients are equal to zero. Use $ \alpha = 0.05$, and be sure to give the critical value of $t$. 

```{r}
t1 <- 1.57/0.34
t2 <- 23.4/11.2
t3 <- -9.5/7.1

t.critical1 <- qt(0.975, df = 13)
t.critical2 <- qt(0.025, df = 13)

if(t1 > t.critical1){
  print("we reject the null hypotheses that coefficients median family income are equal to zero")
}else{
  print("we accept the null hypotheses that coefficients median family income are equal to zero")
}

```

```{r}
if(t2 > t.critical1){
  print("we reject the null hypotheses that coefficients size of house are equal to zero")
}else{
  print("we accept the null hypotheses that coefficients size of house are equal to zero")
}
```
```{r}
if(t3 < t.critical2){
  print("we reject the null hypotheses that coefficients size of lot are equal to zero")
}else{
  print("we accept the null hypotheses that coefficients size of lot are equal to zero")
}
```


### Question 1(g)
What do you conclude from the variance inflation factors (VIFs)? What (if any) modifications would you recommend in light of the VIFs? 

```{r}
vif.lot <- 11.3
if(vif.lot > 5){
  print("the value has the potential multicollinearity problem, we need to remove this value")
}

```


### Question 1(h)
What is the predicted sales price of a house that is 1500 square feet, on a lot 60´×100´, and in a neighborhood where the median family income is $40,000?

```{r}
sales <- 1.57*40000 + 23.4*1500 + 40000
sales
```



## Question 2

This is Exercise 10 in Chapter 9 of the Textbook [R].

### Question 2(a)
Using R and the [Hypothetical UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv), construct a regression equation using housing price as the dependent variable, and bedrooms, bathrooms, date built, garage, fireplace, floor area, and whether the home is detached as the independent variables. Investigate the importance of multicollinearity and outliers. Comment on the weaknesses of this specification, and on the results. 

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df <- read.csv(file = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv", header = T)
y <- df$price
x1 <- df$bedrooms
x2 <- df$bathrooms
x3 <- df$datebuilt
x4 <- df$garage
x5 <- df$fireplace
x6 <- df$floorarea
x7 <- df$detached

m <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7, data = df)
m

```

```{r}
summary(m)
```

```{r}
plot(m)
```

```{r}
result <- anova(m)
r2_gargage <- (result$`Sum Sq`[1])/(sum(result$`Sum Sq`))
VIF_gargage <- 1/(1-r2_gargage)
r2_bedrooms <-(result$`Sum Sq`[2])/(sum(result$`Sum Sq`))
VIF_bedrooms <- 1/(1-r2_bedrooms)
r2_bathrooms <-(result$`Sum Sq`[3])/(sum(result$`Sum Sq`))
VIF_bathrooms <- 1/(1-r2_bathrooms)
r2_datebuilt <-(result$`Sum Sq`[4])/(sum(result$`Sum Sq`))
VIF_datebuilt <- 1/(1-r2_datebuilt)
r2_floorarea <-(result$`Sum Sq`[5])/(sum(result$`Sum Sq`))
VIF_floorarea <- 1/(1-r2_floorarea)
r2_detached <-(result$`Sum Sq`[6])/(sum(result$`Sum Sq`))
VIF_detached <- 1/(1-r2_detached)
r2_fireplace<-(result$`Sum Sq`[7])/(sum(result$`Sum Sq`))
VIF_fireplacee <- 1/(1-r2_fireplace)
c2 <- rbind(VIF_gargage,VIF_bedrooms,VIF_bathrooms,VIF_datebuilt,VIF_floorarea,VIF_detached,VIF_fireplacee)
c1 <- rbind("1","2","3","4","5","6","7")
Vif <- cbind(c1,c2)
colnames(Vif) <- c("Number","VIF")
library(knitr)
kable(Vif)
```


### Question 2(b)
Attempt to improve the regression equation found in (a). Justify your decisions in constructing and carrying out the analysis.

```{r}
full.model <- lm(y ~  x1 + x2 + x3 + x4 + x5 + x6 + x7, data = df)
reduced.model <- step(full.model, direction="backward")
```

```{r}
m2 <- lm(y ~ x2 + x3 + x4 + x5 + x6 + x7, data = df)
m2
anova(m2)
```

```{r}
print("housing price = -261362 + bathrooms*2584 + datebuilt*133.4 + garge*11040.8 + fireplace*(-3127.2) + floorarea*386 + detached*13311.2")
```



## Question 3

This is Exercise 11 in Chapter 9 of the Textbook [R].

### Question 3(a)

Using R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) described in Section 1.9.2, construct a regression equation using housing sales price as the dependent variable, and number of bedrooms, lot size, finished square footage in the house, age of house, and number of bathrooms, as the independent variables. Investigate the importance of multicollinearity and outliers. Comment on the weaknesses of this specification, and on the results. 

```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df1 <- read.csv(file =  "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv", header = T)
y1 <- df1$SalePrice
x1 <- df1$Bedrms
x2 <- df1$LotSize
x3 <- df1$FinSqft
x4 <- df1$Age
x5 <- df1$Baths

m3 <- lm(y1 ~ x1 + x2 + x3 + x4 + x5, data = df1)
m3

```

```{r}
summary(m3)
```

```{r}
plot(m3)
```

```{r}
result1 <- anova(m3)
r2_Bedrms <- (result1$`Sum Sq`[1])/(sum(result1$`Sum Sq`))
VIF_Bedrms <- 1/(1-r2_Bedrms)
r2_LotSize <-(result1$`Sum Sq`[2])/(sum(result1$`Sum Sq`))
VIF_LotSize <- 1/(1-r2_LotSize)
r2_FinSqft <-(result1$`Sum Sq`[3])/(sum(result1$`Sum Sq`))
VIF_FinSqft <- 1/(1-r2_FinSqft)
r2_Age <-(result1$`Sum Sq`[4])/(sum(result1$`Sum Sq`))
VIF_Age <- 1/(1-r2_Age)
r2_Baths <-(result1$`Sum Sq`[5])/(sum(result1$`Sum Sq`))
VIF_Baths <- 1/(1-r2_Baths)
c2 <- rbind(VIF_Bedrms,VIF_LotSize,VIF_FinSqft,VIF_Age,VIF_Baths)
c1 <- rbind("1","2","3","4","5")
Vif <- cbind(c1,c2)
colnames(Vif) <- c("Number","VIF")
library(knitr)
kable(Vif)
```


### Question 3(b)
Attempt to improve the regression equation found in (a). Justify your decisions in constructing and carrying out the analysis.

```{r}
full <- lm(y1 ~  x1 + x2 + x3 + x4 + x5, data = df)
reduced <- step(full, direction="backward")
```

```{r}
summary(reduced)
```

```{r}
print("salePrice = -26240 + Bedrms*(-28760) + Lotsize*3.58 + FinSqft*100.7 + Baths*23020")
```

